Equivalence Testing and the Second Generation P-Value

### 2018-08-22 manuscript.Rmd
initial version papaja manuscript and .bib

---
title: "Equivalence Testing and the Second Generation P-Value"
author: "Daniël Lakens & Marie Delacre"
date: "16 august 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=99)
```
```{r, include=FALSE}
library(TOSTER)
```

```{r, include=FALSE}
source("TOST_to_SGPV.R")
source("p_delta_function.R")

```
The RMarkdown file with the reproducible code of this text is [here](https://github.com/Lakens/TOST_vs_SGPV/blob/master/sgpv_vs_TOST.Rmd).

To test predictions most researchers predominantly rely on null-hypothesis tests. 
This statistical approach can be used to examine whether observed data is sufficiently surprising under the null hypothesis to reject an effect size of zero. 
Null-hypothesis tests have an important limitation, in that this procedure can only reject the hypothesis that there is no effect, while scientists should also be able to provide statistical support for *equivalence*. 
When testing for equivalence researchers aim to examine whether the effect is practically zero, or that there is no meaningful effect. 
By specifying a range around the null hypothesis of values that are deemed practically equivalent to the absence of an effect (i.e., 0 ± 0.3) the observed data can be compared against an *equivalence range* and researchers can test if a meaningful effect is absent (e.g., Hauk & Anderson, 1985; Kruschke, 2018; Metzler, 1974; Rogers, Howard, & Vessey, 1993; Serlin & Lapsey, 1985; Spiegelhalter, Freedman, & Parmar, 1994; Wellek, 2010; Westlake, 1972).

Second generation p-values (SGPV) were recently proposed to as a descriptive statistic that can loosely be interpreted as ‘the proportion of data-supported hypotheses that are also null hypotheses’ (Blume, D’Agostino McGowan, Dupont, & Greevy, 2018). 
The researcher specifies an equivalence range around a classical simple  null hypothesis of parameter values that are considered practically equivalent to the null hypothesis. 
The SGPV measures the degree to which a set of data-supported parameter values falls within the interval null hypothesis. If the estimation interval falls completely within the equivalence range, the SGPV is 1. 
If it lies beyond the equivalence range, the SGPV is 0. Otherwise the SGPV is a value between 0 and 1 that expresses the overlap of data-supported hypotheses and the equivalence range. 
When calculating the SGPV the set of data-supported parameter values can be represented by a confidence interval (CI) or credible intervals or Likelihood support intervals (SI). 
When a confidence interval is used, the SGPV and equivalence tests such as the Two One-Sided Tests (TOST) procedure (Lakens, 2017; Meyners, 2012; Schuirman, 1987) appear to have close ties. 
Here, we aim to examine the similarities and differences between the TOST procedure and the SGPV.

@@

To examine the relation between the TOST p-value and the SGPV we can calculate both statistics across a range of observed effect sizes. In Figure 1 *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for hypothetical one-sample *t*-tests for *all* means that can be observed in studies ranging from 140 to 150 (on the x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), the observed standard deviation is assumed to be 2, and the sample size is 100. For example, for the left-most point in Figure 1 the SGPV and the TOST *p*-value is calculated for a hypothetical study with a sample size of 100, an observed standard deviation of 2, and an observed mean of 140, where the *p*-value for the equivalence test is 1, and the SGPV is 0. Our conclusions about the relationship between TOST *p*-values and SGPV in this article are not dependent upon any specific example.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

lines(sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
#abline(h = c(0.975, 0.5, 0.025), col = "grey", lwd = 1, lty = 3)
```
*Figure 1*: Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2.

The SGPV treats the equivalence range as the null-hypothesis, while the TOST procedure treats the values outside of the equivalence range as the null-hypothesis. For ease of comparison we can reverse the SGPV (by calculating 1-SGPV) to make the values more easily comparable. We see that the *p*-value from the TOST procedure and the SGPV follow each other closely. 
```{r 1-sgpv_tost, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and 1-SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
```
*Figure 2*: Comparison of *p*-values from TOST (black line) and 1-SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2.

When the observed sample mean is 145 and we are testing against equivalence bounds of 143 and 147 using the TOST procedure for a one-sample *t*-test with a sample size of 30 and a standard deviation of 2, the equivalence test is significant, *t*(`r tost_res1$TOST_df`) = `r round(tost_res1$TOST_t1, digits = 2)`, *p* = `r round(tost_res1$TOST_p1, digits =  4)`. Because the 95% CI falls completely within the equivalence bounds, the SGPV is `r SGPV1` (see Figure 1).

On the other hand, if the observed mean is 140, the equivalence test is not significant (the observed mean is far outside the equivalence range of 143 to 147), *t*(`r tost_res2$TOST_df`) = `r round(tost_res2$TOST_t1, digits = 2)`, *p* = `r round(tost_res2$TOST_p1, digits =  4)` (or more accurately, *p* > .999 as *p*-values are bounded between 0 and 1). Because the 95% CI falls completely outside the equivalence bounds, the SGPV is `r SGPV2` (see Figure 1).

@@

It is clear the SGPV and the *p*-value from TOST are closely related. We can think of the SGPV as a straight line that will always overlap the *p*-value from an equivalence test in 3 points. When the TOST *p*-value is 0.5, the SGPV is also 0.5 (note that the reverse is not true). The SGPV is 50% when the observed mean falls exactly on the lower or upper equivalence bound. When the observed mean equals the equivalence bound, the difference between the mean in the data and the equivalence bound is 0, the *t*-value for the equivalence test is also 0, and thus the *p*-value is 0.5 (situation A, Figure 3). 
```{r, include=FALSE}
m <- 146.5
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.025

@@

```
```{r, echo=FALSE, dpi=600, fig.width=5, fig.height=4}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(-3, 5),
     yaxt = "n",
     ylab = "SGPV or TOST p-value",
     xlab = "Mean Difference")
axis(2, at = c(0.25,0.5,0.75), labels = c("C", "B", "A"), las = 1)
axis(1, at = c(-3,-2,-1,0,1,2,3,4,5), las = 1)
abline(v = tost_res1$high_eqbound, 
       lty = 2)
abline(v = tost_res1$low_eqbound, 
       lty = 2)
abline(v = 0, 
       lty = 2, 
       col = "grey")

@@

points(x = tost_res1$diff, 
       y = 0.75, 
       pch = 15, 
       cex = 2)
segments(tost_res1$LL_CI_TOST, 
         0.75, 
         tost_res1$UL_CI_TOST, 
         0.75, 
         lwd = 3)
par(new=TRUE)
curve(dnorm(x, 147.48 - 144.5, 0.5), from=1.5, to=4.5, ylim = c(-2, 6), xlim = c(-3, 5), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 145.520012 - 144.5, 0.5), from=-0.5, to=2.5, ylim = c(-4, 4), xlim = c(-3, 5), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 2, 0.5), from=0.5, to=3.5, ylim = c(-6, 2), xlim = c(-3, 5), xlab = "", ylab = "", axes = FALSE)
```
*Figure 3*: Means, normal distribution, and 95% CI for three example datasets that illustrate the relationship between *p*-values from TOST and SGPV.

Two other points always have to overlap. When the 95% CI falls completely, but only just inside the equivalence region, the TOST (which relies on a one-sided test) should be significant at an alpha level of 0.025. When the SGPV changes from <1 to 1 the 95% CI is exactly equal to one of the equivalence bounds (see situation B in Figure 3, where the 95% CI falls completely inside the equivalence bounds) the TOST *p*-value is 0.025. The third point where the SGPV and the *p*-value from the TOST procedure should overlap is where the SGPV  changes from a positive value (i.e., 0.0001) to 0 (when the 95% CI completely falls outside of the equivalence bound, see situation C in Figure 3). When the 95% CI touches the outside of the equivalence bound and the TOST *p*-value will be 0.975.

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
```{r, echo=FALSE, dpi=600, fig.width=5, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(-3, 3),
     yaxt = "n",
     ylab = "",
     xlab = "Mean Difference")
axis(2, at = c(0.2,0.4,0.6,0.8), labels = c("D", "C", "B", "A"), las = 1)
abline(v = tost_res1$high_eqbound, 
       lty = 2)
abline(v = tost_res1$low_eqbound, 
       lty = 2)
abline(v = 0, 
       lty = 2, 
       col = "grey")

@@

par(new=TRUE)
curve(dnorm(x, 1.2, 0.5), from=-0.3, to=2.7, ylim = c(-2, 8), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.3, 0.5), from=-0.2, to=2.8, ylim = c(-4, 6), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.4, 0.5), from=-0.1, to=2.9, ylim = c(-6, 4), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.5, 0.5), from=0, to=3, ylim = c(-8, 2), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
```
*Figure 4*: Means, normal distribution, and 95% CI for samples where the true population mean is 1.5, 1.4, 1.3, and 1.2.

@@

As we move the means closer to the test value in steps of 0.1 across A to D the *p*-value calculated for normally distributed data is not uniformly distributed. The probability of observing data more extreme than the upper bound of 2 is (from A to D) `r round(tail1, digits = 3)`, `r round(tail2, digits = 3)`, `r round(tail3, digits = 3)`, and `r round(tail4, digits = 3)`. As we can see, the difference between A and B (`r round(tail1-tail2, digits = 3)`) is not the same as the difference between C And D (`r round(tail3-tail4, digits = 3)`). Indeed, the difference in *p*-values is the largest as you start at *p* = 0.5 (when the observed mean falls on the test value), which is why the line in Figure 1 is the steepest at *p* = 0.5. Note that where the SGPV reaches 1 or 0, *p*-values closely approximate 0 and 1, but never reach these values.

##When are the SGPV and Equivalence Test Unrelated?

There are 4 situations where *p*-values from TOST and SGPV are unrelated. The first two situations were discussed earlier, and can be seen in Figure  1. When the SGPV is either 0 or 1 *p*-values from the equivalence test fall between 0.975 and 1 or between 0 and 0.025. Because p-values approach 0 or 1, but are never exactly 0 or 1, while the SGPV is exactly 0 or 1, the two statistics are completely unrelated. The easiest way to see this is by plotting the SGPV against the *p*-value from the TOST procedure. The situations where the SPGV and *p*-values from the TOST procedure are unrelated are indicated by the parts of the curve where there are vertical lines at SGPV of 0 and 1.
```{r, include=FALSE}
step = 0.01

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(sgpv_list, 
     p_tost_list,
     type="l",
     lwd = 3, 
     ylim = c(0, 1), 
     xlim = c(0, 1),
     # yaxt = "n",
     # xaxt = "n",
     ylab = "TOST p-value",
     xlab = "SGPV")
```
*Figure 5*: The relationship between *p*-values from the TOST procedure and the SGPV for the same scenario as in Figure 1.

A third situation in which the SGPV deviates strongly from the TOST *p*-value is whenever the CI is more than twice as wide as the equivalence range, and the CI overlaps with the upper *and* lower equivalence bound. In this situation the normal calculation of the proportion of overlap is skipped, and the SGPV is set to 0.5 instead. Blume et al. (2018) call this the 'small sample correction factor'. However, it is not a correction in the typical sense of the word, since the SGPV is not adjusted to any 'correct' value. When the normal calculation would be 'misleading' (i.e., the SGPV would be small, which normally would suggest support for the alternative hypothesis, when all values in the equivalence range are also supported), the SGPV is set to 0.5 which according to Blume and colleagues signal the SGPV is 'uninformative'. Note that the CI can be twice as wide as the equivalence range whenever the sample size is small (and the confidence interval width is large) *or* when then equivalence range is narrow. It is therefore not so much a 'small sample correction' as it is an exception to the typical calculation of the SGPV whenever the ratio of the confidence interval width to the equivalence range exceeds 2:1 and the CI overlaps with the upper and lower bounds.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 600),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,600,100), labels = seq(-3,3,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

@@

```
*Figure 6*: Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis). Because the sample size is small (n = 10) and the CI is more than twice as wide as the equivalence range (set to -0.4 to 0.4), the SGPV is set to 0.5 (horizontal lightgrey line) across a range of observed means.

>>>>

---
title             : "Equivalence Testing and the Second Generation P-Value"
shorttitle        : "TOST vs. SGPV"

author:
  - name          : "Dani&euml;l Lakens"
    affiliation   : "1"
    corresponding : yes
    address       : "Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands"
    email         : "D.Lakens@tue.nl"
  - name          : "Marie Delacre"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Eindhoven University of Technology, Eindhoven, The Netherlands"
  - id            : "2"
    institution   : "Service of Analysis of the Data, Université Libre de Bruxelles, Belgium"

author_note: |
  All code associated with this article, including the reproducible manuscript, is available from https://github.com/Lakens/TOST_vs_SGPV.

abstract: |
  We aim to compare TOST and SGPV.

keywords          : "equivalence testing, second generation p-values, p-values, hypothesis testing, statistics"
wordcount         : ""

bibliography      : ["TOST_vs_SGPV.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r include = FALSE}
library("TOSTER")
library("papaja")
library("citr")
```
```{r, include=FALSE}
source("TOST_to_SGPV.R")
source("p_delta_function.R")

```

To test predictions most researchers predominantly rely on null-hypothesis tests. 
This statistical approach can be used to examine whether observed data is sufficiently surprising under the null hypothesis to reject an effect size of zero. 
Null-hypothesis tests have an important limitation, in that this procedure can only reject the hypothesis that there is no effect, while scientists should also be able to provide statistical support for *equivalence*. 
When testing for equivalence researchers aim to examine whether the effect is practically zero, or that there is no meaningful effect. 
By specifying a range around the null hypothesis of values that are deemed practically equivalent to the absence of an effect (i.e., 0 ± 0.3) the observed data can be compared against an *equivalence range* and researchers can test if a meaningful effect is absent [@hauck_new_1984;@kruschke_rejecting_2018;@rogers_using_1993;@serlin_rationality_1985;@spiegelhalter_bayesian_1994;@wellek_testing_2010;@westlake_use_1972].

Second generation p-values (SGPV) were recently proposed to as a descriptive statistic that can loosely be interpreted as ‘the proportion of data-supported hypotheses that are also null hypotheses’ [@blume_second-generation_2018]. 
The researcher specifies an equivalence range around a classical simple  null hypothesis of parameter values that are considered practically equivalent to the null hypothesis. 
The SGPV measures the degree to which a set of data-supported parameter values falls within the interval null hypothesis. If the estimation interval falls completely within the equivalence range, the SGPV is 1. 
If it lies beyond the equivalence range, the SGPV is 0. Otherwise the SGPV is a value between 0 and 1 that expresses the overlap of data-supported hypotheses and the equivalence range. 
When calculating the SGPV the set of data-supported parameter values can be represented by a confidence interval (CI) or credible intervals or Likelihood support intervals (SI). 
When a confidence interval is used, the SGPV and equivalence tests such as the Two One-Sided Tests (TOST) procedure [@lakens_equivalence_2017;@meyners_equivalence_2012;@schuirmann_comparison_1987] appear to have close ties. 
Here, we aim to examine the similarities and differences between the TOST procedure and the SGPV.

@@

To examine the relation between the TOST p-value and the SGPV we can calculate both statistics across a range of observed effect sizes. In \ref{fig:TOSTSGPV1} *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for hypothetical one-sample *t*-tests for *all* means that can be observed in studies ranging from 140 to 150 (on the x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), the observed standard deviation is assumed to be 2, and the sample size is 100. For example, for the left-most point in \ref{fig:TOSTSGPV1} the SGPV and the TOST *p*-value is calculated for a hypothetical study with a sample size of 100, an observed standard deviation of 2, and an observed mean of 140, where the *p*-value for the equivalence test is 1, and the SGPV is 0. Our conclusions about the relationship between TOST *p*-values and SGPV in this article are not dependent upon any specific example.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```

```{r, TOSTSGPV1, fig.cap="Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)
lines(sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
```

The SGPV treats the equivalence range as the null-hypothesis, while the TOST procedure treats the values outside of the equivalence range as the null-hypothesis. For ease of comparison we can reverse the SGPV (by calculating 1-SGPV in Figure \\ref{fig:TOSTSGPV2}) to make the values more easily comparable. We see that the *p*-value from the TOST procedure and the SGPV follow each other closely.

```{r, TOSTSGPV2, fig.cap="Comparison of *p*-values from TOST (black line) and 1-SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and 1-SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
```

When the observed sample mean is 145 and we are testing against equivalence bounds of 143 and 147 using the TOST procedure for a one-sample *t*-test with a sample size of 30 and a standard deviation of 2, the equivalence test is significant, *t*(`r tost_res1$TOST_df`) = `r round(tost_res1$TOST_t1, digits = 2)`, *p* = `r round(tost_res1$TOST_p1, digits =  4)`. Because the 95% CI falls completely within the equivalence bounds, the SGPV is `r SGPV1` (see \ref{fig:TOSTSGPV1}).

On the other hand, if the observed mean is 140, the equivalence test is not significant (the observed mean is far outside the equivalence range of 143 to 147), *t*(`r tost_res2$TOST_df`) = `r round(tost_res2$TOST_t1, digits = 2)`, *p* = `r round(tost_res2$TOST_p1, digits =  4)` (or more accurately, *p* > .999 as *p*-values are bounded between 0 and 1). Because the 95% CI falls completely outside the equivalence bounds, the SGPV is `r SGPV2` (see \ref{fig:TOSTSGPV1}).

@@

It is clear the SGPV and the *p*-value from TOST are closely related. We can think of the SGPV as a straight line that will always overlap the *p*-value from an equivalence test in 3 points. When the TOST *p*-value is 0.5, the SGPV is also 0.5 (note that the reverse is not true). The SGPV is 50% when the observed mean falls exactly on the lower or upper equivalence bound. When the observed mean equals the equivalence bound, the difference between the mean in the data and the equivalence bound is 0, the *t*-value for the equivalence test is also 0, and thus the *p*-value is 0.5 (situation A, \ref{fig:TOSTSGPV3}). 
```{r, include=FALSE}
m <- 146.5
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.025

@@

```
```{r, TOSTSGPV3, fig.cap="Means, normal distribution, and 95% CI for three example datasets that illustrate the relationship between *p*-values from TOST and SGPV."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(-3, 5),
     yaxt = "n",
     ylab = "SGPV or TOST p-value",
     xlab = "Mean Difference")
axis(2, at = c(0.25,0.5,0.75), labels = c("C", "B", "A"), las = 1)
axis(1, at = c(-3,-2,-1,0,1,2,3,4,5), las = 1)
abline(v = tost_res1$high_eqbound, 
       lty = 2)
abline(v = tost_res1$low_eqbound, 
       lty = 2)
abline(v = 0, 
       lty = 2, 
       col = "grey")

@@

points(x = tost_res1$diff, 
       y = 0.75, 
       pch = 15, 
       cex = 2)
segments(tost_res1$LL_CI_TOST, 
         0.75, 
         tost_res1$UL_CI_TOST, 
         0.75, 
         lwd = 3)
par(new=TRUE)
curve(dnorm(x, 147.48 - 144.5, 0.5), from=1.5, to=4.5, ylim = c(-2, 6), xlim = c(-3, 5), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 145.520012 - 144.5, 0.5), from=-0.5, to=2.5, ylim = c(-4, 4), xlim = c(-3, 5), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 2, 0.5), from=0.5, to=3.5, ylim = c(-6, 2), xlim = c(-3, 5), xlab = "", ylab = "", axes = FALSE)
```

Two other points always have to overlap. When the 95% CI falls completely, but only just inside the equivalence region, the TOST (which relies on a one-sided test) should be significant at an alpha level of 0.025. When the SGPV changes from <1 to 1 the 95% CI is exactly equal to one of the equivalence bounds (see situation B in \ref{fig:TOSTSGPV3}, where the 95% CI falls completely inside the equivalence bounds) the TOST *p*-value is 0.025. The third point where the SGPV and the *p*-value from the TOST procedure should overlap is where the SGPV  changes from a positive value (i.e., 0.0001) to 0 (when the 95% CI completely falls outside of the equivalence bound, see situation C in \ref{fig:TOSTSGPV3}). When the 95% CI touches the outside of the equivalence bound and the TOST *p*-value will be 0.975.

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
```{r, TOSTSGPV4, fig.cap="Means, normal distribution, and 95% CI for samples where the true population mean is 1.5, 1.4, 1.3, and 1.2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(-3, 3),
     yaxt = "n",
     ylab = "",
     xlab = "Mean Difference")
axis(2, at = c(0.2,0.4,0.6,0.8), labels = c("D", "C", "B", "A"), las = 1)
abline(v = tost_res1$high_eqbound, 
       lty = 2)
abline(v = tost_res1$low_eqbound, 
       lty = 2)
abline(v = 0, 
       lty = 2, 
       col = "grey")

@@

par(new=TRUE)
curve(dnorm(x, 1.2, 0.5), from=-0.3, to=2.7, ylim = c(-2, 8), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.3, 0.5), from=-0.2, to=2.8, ylim = c(-4, 6), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.4, 0.5), from=-0.1, to=2.9, ylim = c(-6, 4), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.5, 0.5), from=0, to=3, ylim = c(-8, 2), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
```

@@

As we move the means closer to the test value in steps of 0.1 across A to D the *p*-value calculated for normally distributed data is not uniformly distributed. The probability of observing data more extreme than the upper bound of 2 is (from A to D) `r round(tail1, digits = 3)`, `r round(tail2, digits = 3)`, `r round(tail3, digits = 3)`, and `r round(tail4, digits = 3)`. As we can see, the difference between A and B (`r round(tail1-tail2, digits = 3)`) is not the same as the difference between C And D (`r round(tail3-tail4, digits = 3)`). Indeed, the difference in *p*-values is the largest as you start at *p* = 0.5 (when the observed mean falls on the test value), which is why the line in \ref{fig:TOSTSGPV1} is the steepest at *p* = 0.5. Note that where the SGPV reaches 1 or 0, *p*-values closely approximate 0 and 1, but never reach these values.
##When are the SGPV and Equivalence Test Unrelated?
There are 4 situations where *p*-values from TOST and SGPV are unrelated. The first two situations were discussed earlier, and can be seen in \ref{fig:TOSTSGPV1}. When the SGPV is either 0 or 1 *p*-values from the equivalence test fall between 0.975 and 1 or between 0 and 0.025. Because p-values approach 0 or 1, but are never exactly 0 or 1, while the SGPV is exactly 0 or 1, the two statistics are completely unrelated. The easiest way to see this is by plotting the SGPV against the *p*-value from the TOST procedure. The situations where the SPGV and *p*-values from the TOST procedure are unrelated are indicated by the parts of the curve where there are vertical lines at SGPV of 0 and 1.
```{r, include=FALSE}
step = 0.01

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, TOSTSGPV5, fig.cap="The relationship between *p*-values from the TOST procedure and the SGPV for the same scenario as in Figure \\ref{fig:TOSTSGPV1}."}
plot(sgpv_list, 
     p_tost_list,
     type="l",
     lwd = 3, 
     ylim = c(0, 1), 
     xlim = c(0, 1),
     # yaxt = "n",
     # xaxt = "n",
     ylab = "TOST p-value",
     xlab = "SGPV")
```

A third situation in which the SGPV deviates strongly from the TOST *p*-value is whenever the CI is more than twice as wide as the equivalence range, and the CI overlaps with the upper *and* lower equivalence bound. In this situation the normal calculation of the proportion of overlap is skipped, and the SGPV is set to 0.5 instead.  @blume_second-generation_2018 call this the 'small sample correction factor'. However, it is not a correction in the typical sense of the word, since the SGPV is not adjusted to any 'correct' value. When the normal calculation would be 'misleading' (i.e., the SGPV would be small, which normally would suggest support for the alternative hypothesis, when all values in the equivalence range are also supported), the SGPV is set to 0.5 which according to Blume and colleagues signal the SGPV is 'uninformative'. Note that the CI can be twice as wide as the equivalence range whenever the sample size is small (and the confidence interval width is large) *or* when then equivalence range is narrow. It is therefore not so much a 'small sample correction' as it is an exception to the typical calculation of the SGPV whenever the ratio of the confidence interval width to the equivalence range exceeds 2:1 and the CI overlaps with the upper and lower bounds.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, TOSTSGPV6, fig.cap="Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis). Because the sample size is small (n = 10) and the CI is more than twice as wide as the equivalence range (set to -0.4 to 0.4), the SGPV is set to 0.5 (horizontal lightgrey line) across a range of observed means."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 600),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,600,100), labels = seq(-3,3,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

@@

```
